### 2023-11-28 paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Latin Modern Roman}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A density ratio framework for evaluating the utility of synthetic data},
  pdfauthor={Thom Benjamin Volker; Peter-Paul de Wolf; Erik-Jan van Kesteren},
  pdfkeywords={TODO, TODO},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\usepackage{setspace}
\doublespacing
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Density ratios for utility }
\title{A density ratio framework for evaluating the utility of synthetic
data}
\def\asep{\\\\\\ } % default: all authors on same column
\def\asep{\And }
\author{\textbf{Thom Benjamin
Volker}~\orcidlink{0000-0002-2408-7820}\\Methodology and Statistics
\textbar{} Methodology\\Utrecht University \textbar{} Statistics
Netherlands\\Utrecht,\ 3584CH\\\href{mailto:t.b.volker@uu.nl}{t.b.volker@uu.nl}\asep\textbf{Peter-Paul
de Wolf}\\Methodology\\Statistics Netherlands\\The
Hague,\ 2490HA\\\href{mailto:pp.dewolf@cbs.nl}{pp.dewolf@cbs.nl}\asep\textbf{Erik-Jan
van Kesteren}\\Methodology and Statistics\\Utrecht
University\\Utrecht,\ 3584CH\\\href{mailto:e.vankesteren1@uu.nl}{e.vankesteren1@uu.nl}}
\date{}
\begin{document}
\maketitle
\begin{abstract}
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
TODO \sep 
TODO

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Openly accessible research data accelerates scientific progress
tremendously {[}ADD REFERENCES{]}. Open data allows third-party
researchers to answer research questions with already collected data,
substantially lowering the costs of research. Sharing data in
combination with code allows others to validate research findings and
build upon the work. Students can benefit from open data, as it fosters
education with realistic data, as well as the general public, through
stimulating citizen science projects. However, openly available research
data is often at odds with privacy constraints on the dissemination of
the data. The information may cause harm to individuals or organizations
if made publicly available. Additionally, sharing their information
openly may withhold them from participating in future research. These
constraints have been named among the biggest hurdles in the advancement
of computational social science
(\protect\hyperlink{ref-lazer_css_2009}{Lazer et al. 2009}), and among
top reasons for companies to not share their data with researchers
(\protect\hyperlink{ref-fpf_2017}{Future of Privacy Forum 2017}). To
overcome these obstacles, data providers can employ a suite of different
disclosure limitation techniques before sharing data, for example
top-coding, record-swapping or adding noise (e.g.,
\protect\hyperlink{ref-hundepool_disclosure_2012}{Hundepool et al.
2012}; \protect\hyperlink{ref-willenborg_elements_2012}{Willenborg and
De Waal 2012}).

Recently, synthetic data has gained substantial traction as a means to
disclosure limitation. National statistical institutes and other
government agencies have adopted the synthetic data framework to
facilitate the use of their data (e.g.,
\protect\hyperlink{ref-SIPP_Beta_2006}{Abowd, Stinson, and Benedetto
2006}; \protect\hyperlink{ref-hawala_synthetic_2008}{Hawala 2008};
\protect\hyperlink{ref-drechsler2012}{Drechsler 2012}). Researchers
started using synthetic data to train machine learning models
(\protect\hyperlink{ref-nikolenko2021}{Nikolenko 2021}), and began to
share synthetic versions of their research data to comply with open
science standards (e.g., \protect\hyperlink{ref-vandewiel2023}{van de
Wiel et al. 2023}; \protect\hyperlink{ref-obermeyer2019}{Obermeyer et
al. 2019}; \protect\hyperlink{ref-zettler2021}{Zettler et al. 2021}).
The idea of synthetic data is to replace some, or all, of the observed
values in a data set by synthetic records that are generated from some
model (e.g., \protect\hyperlink{ref-little_statistical_1993}{Little
1993}; \protect\hyperlink{ref-rubin_statistical_1993}{Rubin 1993};
\protect\hyperlink{ref-drechsler2011synthetic}{Drechsler 2011}). If only
some values are replaced, disclosure risks can be reduced because the
sensitive or identifying values do not correspond to their true values
anymore. If all values are replaced, there is also no one-to-one mapping
between the original and the synthetic data, further reducing the
disclosure risk.

Many different methods have been proposed to generate synthetic data.
Traditionally, these were closely connected to methods used for multiple
imputation of missing data, such as joint modelling {[}misdc2003{]},
sequential regressions (\protect\hyperlink{ref-nowok2016}{Nowok, Raab,
and Dibben 2016}) or fully conditional specification
(\protect\hyperlink{ref-drechsler2011empirical}{Drechsler and Reiter
2011}; \protect\hyperlink{ref-volker2021}{Volker and Vink 2021}). The
flexibility of sequential regressions and fully conditional
specification is commonly combined with non-parametric imputation
models, such as classification and regression trees
(\protect\hyperlink{ref-reiter2005}{Reiter 2005}), random forests
(\protect\hyperlink{ref-caiola2010}{Caiola and Reiter 2010}) or support
vector machines (\protect\hyperlink{ref-drechsler2010}{Drechsler 2010}),
to avoid distributional assumptions and easily model non-linear
relationships. Recently, significant improvements in generative
modelling sparked the scientific interest in synthetic data in the
computer science community, leading to novel synthesis methods (e.g.,
\protect\hyperlink{ref-patki2016}{Patki, Wedge, and Veeramachaneni
2016}; \protect\hyperlink{ref-xu_ctgan_2019}{Xu et al. 2019}). Combined
with work on formal privacy guarantees such as differential privacy,
this resulted in new models that explicitly control the level of privacy
risk in synthesis methods
(\protect\hyperlink{ref-jordon2018pategan}{Jordon, Yoon, and Schaar
2019}; \protect\hyperlink{ref-Torkzadehmahani2019}{Torkzadehmahani,
Kairouz, and Paten 2019}).

\textbf{\emph{TODO}}

All methods for statistical disclosure limitation alter the data before
these are provided to the public. By doing so, the utility of the
provided data is always lower than the utility of the original data,
because some of the information in the data is sacrificed to protect the
privacy of the respondents. The questions that naturally arise are how
much information in the original data is actually sacrificed, and how
useful the provided data are? Answering this question allows researchers
to decide what the altered data can and cannot be used for, and to
evaluate the worth of conclusions drawn on the basis of these data.
After all, inferences from the altered data are valid only up to the
extent that the perturbation methods approximate the true
data-generating mechanism. For data providers, a detailed assessment of
the quality of the altered data can guide the procedure of altering the
data. Statistical disclosure limitation is often an iterative process:
some disclosure limitation technique is applied on the data, after which
the result is investigated and modifications are made to applied
process. Good measures of data quality are essential to determine the
appropriate mechanisms used to protect the data, and can help to improve
the utility of the data that will be disseminated.

In the statistical disclosure control literature, two different branches
of utility measures have been distinguished: specific utility measures
and general utility measures. \emph{Add one/two sentences on the merit
of visualization when assessing utility of altered data.} Specific
utility measures focus on similarity of results obtained from analyses
performed on the altered data and the original data. For example, after
fitting the same analysis model on both data sets, one can calculate the
confidence interval overlap of the estimated parameters
(\protect\hyperlink{ref-karr_utility_2006}{Karr et al. 2006}).
Alternative measures are ellipsoidal overlap
(\protect\hyperlink{ref-karr_utility_2006}{Karr et al. 2006}), which
extends to confidence interval overlap to a measure that addresses the
joint distribution of all model parameters simultaneous, the
standardized absolute difference between estimates
(\protect\hyperlink{ref-snoke_utility_2018}{Snoke et al. 2018}), and the
ratio of estimates for tabular count data
(\protect\hyperlink{ref-taub2020impact}{\textbf{taub2020impact?}}). As
these measures quantify similarity between estimates from analyses
performed on the observed and altered data, they are informed only to
the extent that data users will recreate those analyses. This can be
highly useful if the data is provided for reproducibility purposes
(e.g., for third parties to evaluate analysis scripts). However, the
goal of distributing the protected data is often to allow researchers to
do novel research with the data. In many practical situations, data
providers thus have have only limited knowledge on the analyses that
will be performed with the altered data. Covering the entire set of
potentially relevant analyses is therefore not feasible. If it was, the
data providers could simply report the (potentially privacy-protected)
results of those analyses performed on the real data, so that access to
the (perturbed) data no longer yields additional benefits (for a similar
argument, see \protect\hyperlink{ref-drechsler_utility_2022}{Drechsler
2022}). Additionally, similarity between results on the analyses that
have been performed gives no guarantee that the results will also be
similar for other analyses. Hence, when determining how useful the
altered data is for novel research, specific utility measures are only
of limited use.

General utility measures attempt to capture how similar the multivariate
distributions of the observed and altered data are. This can be done by,
for instance, estimating the Kullback-Leibler divergence between the
distributions of the observed and altered data
(\protect\hyperlink{ref-karr_utility_2006}{Karr et al. 2006}). An
alternative strategy is to try to discriminate between the observed and
altered data, as is done with the \(pMSE\)
(\protect\hyperlink{ref-snoke_utility_2018}{Snoke et al. 2018};
\protect\hyperlink{ref-woo_utility_2009}{\textbf{woo\_utility\_2009?}}).
In essence, the \(pMSE\) quantifies how well one can predict whether
observations are from the observed or the altered data. If better one
can do this, the more pronounced the differences between the observed
and altered data ought to be. However, various authors have criticized
general utility measures for being too broad. That is, important
discrepancies between the real and altered data might be missed, and an
altered data set that is good in general (i.e., has high global utility)
might still provide results that are far from the truth for some
analyses (see, e.g.,
\protect\hyperlink{ref-drechsler_utility_2022}{Drechsler 2022}).
Additionally, it is not straightforward to determine which prediction
model to use for calculating the \(pMSE\). Specifying a good prediction
model in itself may be a challenging task, especially when the number of
variables is large. When good models are available, different models, or
even different choices of hyperparameters, may yield different results,
potentially rendering ambiguity with respect to which altered data set
is best. Lastly, the output of global utility measures can be hard to
interpret, and say little about the regions in which the synthetic data
do not resemble the true data accurately enough. That is, they give
little guidance on how the quality of the altered data can be improved.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Section 6: our contribution}

\emph{Moet nog verder uitgewerkt worden}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We introduce density ratio estimation to the field of statistical
  disclosure control. Short remark on the idea that density ratio
  estimation is a complicated endeavor, especially if the goal is to
  compare distinct densities. Having to estimate just a single density
  (ratio) is generally much easier.
\item
  Note that density ratio estimation can capture specific and general
  utility measures into a common framework by being applicable on the
  level of the entire data, but also on the subset of variables that is
  relevant in an analysis. Additionally, note that confidence interval
  overlap, ellipsoidal overlap, but also \(pMSE\) and Kullback-Leibler
  divergence, are closely related to density ratio estimation, and can
  be considered from this perspective.
\item
  Create a new utility metric based on density ratio estimation
  (probability with respect to some reference distribution as in
  permutation testing).
\item
  Because density ratio estimation can be difficult when there are many
  variables, we use dimension reduction techniques to capture most of
  the variability in the data in fewer dimensions on which density ratio
  estimation can be applied. A by-product of this is that the
  lower-dimensional subspace allows to create visualizations of
  deviations from the observed data.
\item
  Perform a simulation study to give indications about which methods to
  use (think about how to do this).
\item
  Implement all this in an R-package
\end{enumerate}

\textbf{Section 6: outline article}

In the next section, we describe density ratio estimation and discuss
how this method can be used as to measure utility. Subsequently, we
provide multiple examples that show how density ratio estimation works
in the context of evaluating the quality of synthetic data. Hereafter,
we show in multiple simulations that the method is superior (HOPEFULLY)
to current global utility measures as the \(pMSE\). Lastly, we discuss
the advantages and disadvantages of density ratio estimation as a
utility measure.

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

@Peter-Paul: Eventueel een korte beschrijving van data perturbation
techniques/synthetic data generation hier. Denk je dat dit wat toevoegt
hier?

\textbf{Section 1: density ratio estimation}

In essence, the goal of utility measures is to quantify the similarity
between the multivariate distribution of the observed data with the
distribution of the altered data. If the used data perturbation
techniques, or synthetic data generation models, approximate the
distribution of the real data sufficiently, these distributions should
be highly similar, and analyses on the two data sets should give similar
results. However, estimating the probability distribution of a data set
is known to be one of the most complicated challenges in statistics
{[}E.G. Vapnik 1998{]}. Estimating the probability distribution for both
observed and altered data can lead to errors in both, artificially
magnifying discrepancies between the two. Hence, subsequent comparisons
will be affected by these errors. The procedure can be simplified by
using density ratio estimation, because this only requires to estimate a
single density.

Introduce density ratio estimation as a utility measure. What does this
measure mean/how to interpret it. How to make decisions based on this
measure.

Say something on whether (and if so, how) categorical variables can be
incorporated as well.

\textbf{Section 3: theoretical comparison with conventional approaches
for general utility assessment}

Relate density ratio estimation to specific and general utility
measures. Pick one/two specific utility measures and relate these to
density ratio estimation (ratio of estimates seems straightforward, as
well as confidence interval overlap and ellipsoidal overlap).

Relate density ratio estimation to \(pMSE\) and KL divergence (to some
extent, both are generalizations of density ratio estimation, or at
least are conceptually similar). Give some more information on the
\(pMSE\), describe what it shortcomings are. The quality of the \(pMSE\)
highly depends on the model used to calculate the propensity scores.
Perhaps give an example of logistic regression, which basically
estimates whether the conditional mean of the observed and altered data
is the same. Explain how density ratio estimation can overcome the
shortcomings of the previously mentioned methods.

\textbf{Section 4: Dimension reduction for utility}

The difficulty of density ratio estimation increases with the
dimensionality of the data. Therefore, we follow previous
recommendations to incorporate dimensionality reduction techniques in
density ratio estimation.

Shortly name examples of dimensionality reduction techniques (i.e., PCA;
LFDA or UMAP).

A useful by-product of dimension reduction is that it allows to create
visualizations, and these visualizations can be used to get more insight
in discrepancies between observed and altered data. Show what such
visualizations can look like, and how they can help.

\hypertarget{simulations}{%
\section{Simulations}\label{simulations}}

\hypertarget{small-illustration-example-with-multivariate-gaussian-distributions.}{%
\subsection{Small illustration / example with multivariate Gaussian
distributions.}\label{small-illustration-example-with-multivariate-gaussian-distributions.}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Simple, multivariate normal simulation (e.g., two correlation
  structures, two sample sizes, so \(2 \times 2\) full factorial
  design); basically similar to what we did already.
\end{enumerate}

\hypertarget{more-complex-simulation-more-variables-non-linearities-perhaps-using-real-data.}{%
\subsection{More complex simulation, more variables, non-linearities,
perhaps using real
data.}\label{more-complex-simulation-more-variables-non-linearities-perhaps-using-real-data.}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  More advanced simulation (e.g., some non-linearities, different sample
  sizes)
\end{enumerate}

Have to think about this in more detail still.

\hypertarget{real-data-example}{%
\section{Real data example}\label{real-data-example}}

Clinical records heart-failure data? Misschien ook niet, nog over
nadenken.

Exemplify how utility measures could (should!) be used to improve the
quality of the altered data (e.g., illustrate how models can be adjusted
iteratively based on utility assessment).

\emph{Some notes to self}

Current ways to assess the utility?

\begin{itemize}
\item
  pMSE - logistic, regression, CART models (Snoke, Raab, Nowok, Dibben
  \& Slavkovic, 2018; General and specific utility measures for
  synthetic data AND Woo, Reiter, Oganian \& Karr, 2009; Global measures
  of data utlity for microdata masked for disclosure limitation)
\item
  Kullback-Leiber divergence (Karr, Kohnen, Oganian, Reiter \& Sanil,
  2006; A framework for evaluating the utility of data altered to
  protect confidentiality).
\item
  According to multiple authors, both specific and general utility
  measures have important drawbacks (see Drechsler Utility PSD; cites
  others). Narrow measures potentially focus on analyses that are not
  relevant for the end user, and do not generalize to the analyses that
  are relevant. Global utility measures are generally too broad, and
  important deviations in the synthetic data might be missed. Moreover,
  the measures are typically hard to interpret.
\item
  See Drechsler for a paragraph on fit for purpose measures, that lie
  between general and specific utility measures (i.e., plausibility
  checks such as non-negativity; goodness of fit measures as \(\chi^2\)
  for cross-tabulations; Kolmogorov-Smirnov).
\item
  Drechsler also illustrates that the standardized \(pMSE\) has
  substantial flaws, as the results are highly dependent on the model
  used to estimate the propensity scores, and unable to detect important
  differences in the utility for most of the model specifications.
  Hence, it is claimed that a thorough assessment of utility is
  required.
\end{itemize}

Things to add in new version: - High dimensional example - Expansion of
discussion points: empirical example with weighted analyses individual
data utility - Privacy remark? - When does it not work?

\hypertarget{methodology-1}{%
\section{Methodology}\label{methodology-1}}

TO DO

\hypertarget{simulations-1}{%
\section{Simulations}\label{simulations-1}}

TO DO

\hypertarget{real-data-example-1}{%
\section{Real data example}\label{real-data-example-1}}

TO DO

\hypertarget{results}{%
\section{Results}\label{results}}

TO DO

\hypertarget{discussion-and-conclusion}{%
\section{Discussion and conclusion}\label{discussion-and-conclusion}}

TO DO

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-SIPP_Beta_2006}{}}%
Abowd, John M., Martha Stinson, and Gary Benedetto. 2006. {``Final
Report to the Social Security Administration on the {SIPP/SSA/IRS}
Public Use File Project.''} Longitudinal Employer-Household Dynamics
Program, U.S. Bureau of the Census, Washington, DC.
\url{https://ecommons.cornell.edu/bitstream/handle/1813/43929/SSAfinal.pdf?sequence=3\&isAllowed=y}.

\leavevmode\vadjust pre{\hypertarget{ref-caiola2010}{}}%
Caiola, Gregory, and Jerome P. Reiter. 2010. {``Random Forests for
Generating Partially Synthetic, Categorical Data.''} \emph{Transactions
on Data Privacy} 3: 27--42.
\url{https://doi.org/10.5555/1747335.1747337}.

\leavevmode\vadjust pre{\hypertarget{ref-drechsler2010}{}}%
Drechsler, Jörg. 2010. {``Using Support Vector Machines for Generating
Synthetic Datasets.''} In \emph{Privacy in Statistical Databases},
edited by Josep Domingo-Ferrer and Emmanouil Magkos, 148--61. Berlin,
Heidelberg: Springer Berlin Heidelberg.
\url{https://doi.org/10.1007/978-3-642-15838-4_14}.

\leavevmode\vadjust pre{\hypertarget{ref-drechsler2011synthetic}{}}%
---------. 2011. \emph{Synthetic Datasets for Statistical Disclosure
Control: Theory and Implementation}. New York: Springer Science \&
Business Media. \url{https://doi.org/10.1007/978-1-4614-0326-5}.

\leavevmode\vadjust pre{\hypertarget{ref-drechsler2012}{}}%
---------. 2012. {``New Data Dissemination Approaches in Old Europe
{\textendash} Synthetic Datasets for a German Establishment Survey.''}
\emph{Journal of Applied Statistics} 39 (2): 243--65.
\url{https://doi.org/10.1080/02664763.2011.584523}.

\leavevmode\vadjust pre{\hypertarget{ref-drechsler_utility_2022}{}}%
---------. 2022. {``Challenges in Measuring Utility for Fully Synthetic
Data.''} In \emph{Privacy in Statistical Databases}, edited by Josep
Domingo-Ferrer and Maryline Laurent, 220--33. Cham: Springer
International Publishing.
\url{https://doi.org/10.1007/978-3-031-13945-1_16}.

\leavevmode\vadjust pre{\hypertarget{ref-drechsler2011empirical}{}}%
Drechsler, Jörg, and Jerome P Reiter. 2011. {``An Empirical Evaluation
of Easily Implemented, Nonparametric Methods for Generating Synthetic
Datasets.''} \emph{Computational Statistics \& Data Analysis} 55 (12):
3232--43.

\leavevmode\vadjust pre{\hypertarget{ref-fpf_2017}{}}%
Future of Privacy Forum. 2017. {``Understanding Corporate Data Sharing
Decisions: Practices, Challenges, and Opportunities for Sharing
Corporate Data with Researchers.''}

\leavevmode\vadjust pre{\hypertarget{ref-hawala_synthetic_2008}{}}%
Hawala, Sam. 2008. \emph{Producing Partially Synthetic Data to Avoid
Disclosure}.
\url{http://www.asasrms.org/Proceedings/y2008/Files/301018.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-hundepool_disclosure_2012}{}}%
Hundepool, Anco, Josep Domingo-Ferrer, Luisa Franconi, Sarah Giessing,
Eric Schulte Nordholt, Keith Spicer, and Peter-Paul De Wolf. 2012.
\emph{Statistical Disclosure Control}. John Wiley \& Sons.
\url{https://doi.org/10.1002/9781118348239}.

\leavevmode\vadjust pre{\hypertarget{ref-jordon2018pategan}{}}%
Jordon, James, Jinsung Yoon, and Mihaela van der Schaar. 2019.
{``{PATE}-{GAN}: Generating Synthetic Data with Differential Privacy
Guarantees.''} In \emph{International Conference on Learning
Representations}. \url{https://openreview.net/forum?id=S1zk9iRqF7}.

\leavevmode\vadjust pre{\hypertarget{ref-karr_utility_2006}{}}%
Karr, Alan F., Christine N. Kohnen, Anna Oganian, Jerome P. Reiter, and
Ashish P. Sanil. 2006. {``A Framework for Evaluating the Utility of Data
Altered to Protect Confidentiality.''} \emph{The American Statistician}
60 (3): 224--32. \url{https://doi.org/10.1198/000313006X124640}.

\leavevmode\vadjust pre{\hypertarget{ref-lazer_css_2009}{}}%
Lazer, David, Alex Pentland, Lada Adamic, Sinan Aral, Albert-László
Barabási, Devon Brewer, Nicholas Christakis, et al. 2009.
{``Computational Social Science.''} \emph{Science} 323 (5915): 721--23.
\url{https://doi.org/10.1126/science.1167742}.

\leavevmode\vadjust pre{\hypertarget{ref-little_statistical_1993}{}}%
Little, Roderick J. A. 1993. {``Statistical Analysis of Masked Data.''}
\emph{Journal of Official Statistics} 9 (2): 407--7.
\url{https://www.scb.se/contentassets/ca21efb41fee47d293bbee5bf7be7fb3/statistical-analysis-of-masked-data.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-nikolenko2021}{}}%
Nikolenko, Sergey I. 2021. \emph{Synthetic Data for Deep Learning}.
Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-75178-4}.

\leavevmode\vadjust pre{\hypertarget{ref-nowok2016}{}}%
Nowok, Beata, Gillian M. Raab, and Chris Dibben. 2016.
{``{\emph{Synthpop}:} Bespoke Creation of Synthetic Data in
{\emph{R}}.''} \emph{Journal of Statistical Software} 74 (11).
\url{https://doi.org/10.18637/jss.v074.i11}.

\leavevmode\vadjust pre{\hypertarget{ref-obermeyer2019}{}}%
Obermeyer, Ziad, Brian Powers, Christine Vogeli, and Sendhil
Mullainathan. 2019. {``Dissecting Racial Bias in an Algorithm Used to
Manage the Health of Populations.''} \emph{Science} 366 (6464): 447--53.
\url{https://doi.org/10.1126/science.aax2342}.

\leavevmode\vadjust pre{\hypertarget{ref-patki2016}{}}%
Patki, Neha, Roy Wedge, and Kalyan Veeramachaneni. 2016. {``The
Synthetic Data Vault.''} \emph{2016 IEEE International Conference on
Data Science and Advanced Analytics (DSAA)}, October.
\url{https://doi.org/10.1109/dsaa.2016.49}.

\leavevmode\vadjust pre{\hypertarget{ref-reiter2005}{}}%
Reiter, Jerome P. 2005. {``Using CART to Generate Partially Synthetic
Public Use Microdata.''} \emph{Journal of Official Statistics} 21 (3):
441--62.
\url{https://www.scb.se/contentassets/ca21efb41fee47d293bbee5bf7be7fb3/using-cart-to-generate-partially-synthetic-public-use-microdata.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-rubin_statistical_1993}{}}%
Rubin, Donald B. 1993. {``Statistical Disclosure Limitation.''}
\emph{Journal of Official Statistics} 9 (2): 461--68.
\url{https://www.scb.se/contentassets/ca21efb41fee47d293bbee5bf7be7fb3/discussion-statistical-disclosure-limitation2.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-snoke_utility_2018}{}}%
Snoke, Joshua, Gillian M. Raab, Beata Nowok, Chris Dibben, and
Aleksandra Slavkovic. 2018. {``General and Specific Utility Measures for
Synthetic Data.''} \emph{Journal of the Royal Statistical Society.
Series A (Statistics in Society)} 181 (3): pp. 663--688.
\url{https://doi.org/10.1111/rssa.12358}.

\leavevmode\vadjust pre{\hypertarget{ref-Torkzadehmahani2019}{}}%
Torkzadehmahani, Reihaneh, Peter Kairouz, and Benedict Paten. 2019.
{``{DP-CGAN}: Differentially Private Synthetic Data and Label
Generation.''} In \emph{Proceedings of the IEEE/CVF Conference on
Computer Vision and Pattern Recognition (CVPR) Workshops}.
\url{https://doi.org/10.1109/cvprw.2019.00018}.

\leavevmode\vadjust pre{\hypertarget{ref-vandewiel2023}{}}%
van de Wiel, Mark A., Gwenaël G. R. Leday, Jeroen Hoogland, Martijn W.
Heymans, Erik W. van Zwet, and Ailko H. Zwinderman. 2023. {``Think
Before You Shrink: Alternatives to Default Shrinkage Methods Can Improve
Prediction Accuracy, Calibration and Coverage.''}
\url{https://doi.org/10.48550/ARXIV.2301.09890}.

\leavevmode\vadjust pre{\hypertarget{ref-volker2021}{}}%
Volker, Thom Benjamin, and Gerko Vink. 2021. {``Anonymiced Shareable
Data: Using Mice to Create and Analyze Multiply Imputed Synthetic
Datasets.''} \emph{Psych} 3 (4): 703--16.
\url{https://doi.org/10.3390/psych3040045}.

\leavevmode\vadjust pre{\hypertarget{ref-willenborg_elements_2012}{}}%
Willenborg, Leon, and Ton De Waal. 2012. \emph{Elements of Statistical
Disclosure Control}. Springer Science \& Business Media.
\url{https://doi.org/10.1007/978-1-4613-0121-9}.

\leavevmode\vadjust pre{\hypertarget{ref-xu_ctgan_2019}{}}%
Xu, Lei, Maria Skoularidou, Alfredo Cuesta-Infante, and Kalyan
Veeramachaneni. 2019. {``Modeling Tabular Data Using Conditional GAN.''}
In \emph{Advances in Neural Information Processing Systems}, edited by
H. Wallach, H. Larochelle, A. Beygelzimer, F. dAlché-Buc, E. Fox, and R.
Garnett. Vol. 32. Curran Associates, Inc.
\url{https://proceedings.neurips.cc/paper_files/paper/2019/file/254ed7d2de3b23ab10936522dd547b78-Paper.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-zettler2021}{}}%
Zettler, Ingo, Christoph Schild, Lau Lilleholt, Lara Kroencke, Till
Utesch, Morten Moshagen, Robert Böhm, Mitja D. Back, and Katharina
Geukes. 2021. {``The Role of Personality in COVID-19-Related
Perceptions, Evaluations, and Behaviors: Findings Across Five Samples,
Nine Traits, and 17 Criteria.''} \emph{Social Psychological and
Personality Science} 13 (1): 299--310.
\url{https://doi.org/10.1177/19485506211001680}.

\end{CSLReferences}

\end{document}